---
title: "Logistic Regression Model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remove(list = ls())
library(car)
library(pROC)
library(ComplexHeatmap)
library(circlize)
library(grid)
```
# run estimation.R first
# load the proproceesed data
```{r}
train <- read.csv("../data/derived/df_v1_TRAIN.csv")
test <- read.csv("../data/derived/df_v1_TEST.csv")
## build the train dataset
train %<>%
  dplyr::select(-X) %>%
  mutate(y_mult = factor(num),
         y_bin = factor(ifelse(num == "0", 0, 1)),
         across(c(sex,cp, fbs, restecg, exang, slope, ca, thal), as.factor)
  )
## build the test dataset
test %<>%
  dplyr::select(-X) %>%
  mutate(y_mult = factor(num),
         y_bin = factor(ifelse(num == "0", 0, 1)),
         across(c(sex,cp, fbs, restecg, exang, slope, ca, thal), as.factor)
  )

```

# LR: use the function in estimation on train and test data
## heatmap for training
```{r}
# ********source("../code/estimation.R")*********************
# can't directly use source(), so need manually run this code file
lr_formula = y_bin ~ age+sex+cp+trestbps+chol+fbs+restecg+thalach+exang+oldpeak+slope+ca+thal
lr_fit = glm_v2(lr_formula, df = train, method = "lr")
## train matrix
lr_train_x <- model.matrix(lr_formula, data = train)
colnames(lr_train_x)[colnames(lr_train_x) == "(Intercept)"] <- "Intercept"
lr_train_x <- lr_train_x[, names(lr_fit$beta)]
lr_train_y <- train[,ncol(train)]
## predict probabilities
lr_eta_train <- lr_train_x %*% lr_fit$beta
lr_p_train <- 1 / (1 + exp(-lr_eta_train))
lr_pred_train <- ifelse(lr_p_train >= 0.5, 1, 0) 
## Confusion matrix
lr_conf_train <- table("Prediction" = lr_pred_train, "True Label" = lr_train_y) |> as.matrix()
## Normalize (row-wise proportion: prediction rows)
lr_conf_train_prop <- t(apply(lr_conf_train, 1, function(i) i / colSums(lr_conf_train)))
Heatmap(lr_conf_train_prop, 
        col = colorRamp2(c(0, 1), c("white", "red")),
        name = "Proportion",
        cluster_rows = FALSE, cluster_columns = FALSE,
        row_title = "Prediction", row_names_side = "left",
        column_title = "True Label", column_names_side = "top", column_names_rot = 0,
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.2f", lr_conf_train_prop[i, j]), x, y, gp = gpar(fontsize = 10))
        })

```
## heatmap for testing
```{r}
## test matrix
lr_test_x <- model.matrix(lr_formula, data = test)
colnames(lr_test_x)[colnames(lr_test_x) == "(Intercept)"] <- "Intercept"
lr_test_x <- lr_test_x[, names(lr_fit$beta)]
lr_test_y <- test[,ncol(test)]
## predict probabilities
lr_eta_test <- lr_test_x %*% lr_fit$beta
lr_p_test <- 1 / (1 + exp(-lr_eta_test))
lr_pred_test <- ifelse(lr_p_test >= 0.5, 1, 0) 
## Confusion matrix
lr_conf_test <- table("Prediction" = lr_pred_test, "True Label" = lr_test_y) |> as.matrix()
## Normalize (row-wise proportion: prediction rows)
lr_conf_test_prop <- t(apply(lr_conf_test, 1, function(i) i / colSums(lr_conf_test)))
Heatmap(lr_conf_test_prop, 
        col = colorRamp2(c(0, 1), c("white", "red")),
        name = "Proportion",
        cluster_rows = FALSE, cluster_columns = FALSE,
        row_title = "Prediction", row_names_side = "left",
        column_title = "True Label", column_names_side = "top", column_names_rot = 0,
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.2f", lr_conf_test_prop[i, j]), x, y, gp = gpar(fontsize = 10))
        })

## evaluate performance
roc_obj <- roc(lr_test_y, as.numeric(lr_pred_test))
auc_value <- auc(roc_obj)
print(auc_value)
```

# POLR: use the function in estimation on train and test data
## heatmap for training
```{r}
polr_formula = y_mult ~ age+sex+cp+trestbps+chol+fbs+restecg+thalach+exang+oldpeak+slope+ca+thal
polr_fit = glm_v2(polr_formula, df = train, method = "polr")
## =================plot heatmap for training====================
## Create X_train
polr_train_x <- model.matrix(polr_formula, data = train)[, -1] 
polr_train_x <- polr_train_x[, names(polr_fit$beta)] # Reorder columns
## Linear predictor
polr_eta_train <- polr_train_x %*% polr_fit$beta
## Predict class
n_class <- length(polr_fit$theta) + 1
### No labels, get default levels 1,2,3,4,5
polr_temp_pred_train <- cut(as.numeric(polr_eta_train),
                 breaks = c(-Inf, polr_fit$theta, Inf),
                 right = TRUE)  # gives levels 1,2,3,4,5
### Manually subtract 1 to match 0–4
polr_train_pred <- as.numeric(polr_temp_pred_train) - 1
## True labels for training
polr_train_y <- train$y_mult
## Confusion matrix
polr_train_table <- table("prediction" = polr_train_pred, "true label" = polr_train_y) |> as.matrix()
## Normalize to proportions
polr_train_table_prop <- t(apply(polr_train_table, 1, function(i) i / colSums(polr_train_table)))
## Heatmap for training set
Heatmap(polr_train_table_prop, 
        col = colorRamp2(c(0, 1), c("white", "red")),
        name = "Proportion",
        cluster_rows = FALSE, cluster_columns = FALSE,
        row_title = "Prediction", row_names_side = "left",
        column_title = "True Label", column_names_side = "top", column_names_rot = 0,
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.2f", polr_train_table_prop[i, j]), x, y, gp = gpar(fontsize = 10))
        })

```

## heatmap for testing
```{r}
## Create X_test
polr_test_x <- model.matrix(polr_formula, data = test)[, -1] 
polr_test_x <- polr_test_x[, names(polr_fit$beta)] # Reorder columns
## Linear predictor
polr_eta_test <- polr_test_x %*% polr_fit$beta
### No labels, get default levels 1,2,3,4,5
polr_temp_pred_test <- cut(as.numeric(polr_eta_test),
                 breaks = c(-Inf, polr_fit$theta, Inf),
                 right = TRUE)  # gives levels 1,2,3,4,5
### Manually subtract 1 to match 0–4
polr_test_pred <- as.numeric(polr_temp_pred_test) - 1
## True labels for testing
polr_test_y <- test$y_mult
## Confusion matrix
polr_test_table <- table("prediction" = polr_test_pred, "true label" = polr_test_y) |> as.matrix()
## Normalize to proportions
polr_test_table_prop <- t(apply(polr_test_table, 1, function(i) i / colSums(polr_test_table)))
## Heatmap for testing set
Heatmap(polr_test_table_prop, 
        col = colorRamp2(c(0, 1), c("white", "red")),
        name = "Proportion",
        cluster_rows = FALSE, cluster_columns = FALSE,
        row_title = "Prediction", row_names_side = "left",
        column_title = "True Label", column_names_side = "top", column_names_rot = 0,
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.2f", polr_test_table_prop[i, j]), x, y, gp = gpar(fontsize = 10))
        })

```
max(polr_eta_test) < polr_fit$theta, theta34 is too high, so there's no class4

## heatmaps for binary training
```{r}
polr_train_y_numeric <- as.numeric(as.character(polr_train_y))
polr_train_pred_numeric <- as.numeric(as.character(polr_train_pred))
polr_train_y_binary <- ifelse(polr_train_y_numeric > 0, 1, 0)
polr_train_pred_binary <- ifelse(polr_train_pred_numeric > 0, 1, 0)
polr_train_binary_table <- table("prediction" = polr_train_pred_binary, "true label" = polr_train_y_binary) |> as.matrix()
polr_train_binary_table_prop <- t(apply(polr_train_binary_table, 1, function(i) i / colSums(polr_train_binary_table)))

Heatmap(polr_train_binary_table_prop, 
        col = colorRamp2(c(0, 1), c("white", "red")),
        name = "Proportion",
        cluster_rows = FALSE, cluster_columns = FALSE,
        row_title = "Prediction (Binary)", row_names_side = "left",
        column_title = "True Label (Binary)", column_names_side = "top", column_names_rot = 0,
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.2f", polr_train_binary_table_prop[i, j]), x, y, gp = gpar(fontsize = 10))
        })
```
## heatmaps for binary testing
```{r}
polr_test_y_numeric <- as.numeric(as.character(polr_test_y))
polr_test_pred_numeric <- as.numeric(as.character(polr_test_pred))
polr_test_y_binary <- ifelse(polr_test_y_numeric > 0, 1, 0)
polr_test_pred_binary <- ifelse(polr_test_pred_numeric > 0, 1, 0)

polr_test_binary_table <- table("prediction" = polr_test_pred_binary, "true label" = polr_test_y_binary) |> as.matrix()
polr_test_binary_table_prop <- t(apply(polr_test_binary_table, 1, function(i) i / colSums(polr_test_binary_table)))

Heatmap(polr_test_binary_table_prop, 
        col = colorRamp2(c(0, 1), c("white", "red")),
        name = "Proportion",
        cluster_rows = FALSE, cluster_columns = FALSE,
        row_title = "Prediction (Binary)", row_names_side = "left",
        column_title = "True Label (Binary)", column_names_side = "top", column_names_rot = 0,
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.2f", polr_test_binary_table_prop[i, j]), x, y, gp = gpar(fontsize = 10))
        })

```

